A comprehensive analysis of natural sequence variation within the <i>Arabidopsis

### 2018-07-03 manuscript/manuscript.Rmd
- figure templates made - trees built - diversity across collapsed alignment for A ARFs - initial revision of text -- still need to cut and consolidate

```{r setup}
knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')

#devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
library(r1001genomes)
library(DT)
library(tidyverse)
library(magrittr)
library(RColorBrewer)
library(ggpmisc)
library(ggthemes)
library(ggseqlogo)
library(ggtree)
library(ape)
library(reticulate)
library(scales)
library(viridis)
use_python(python = "/usr/local/bin/python")
```
<!--This is a CSS (Cascading Style Sheet).
 It controls how panels and images are formatted.></!-->
<style>
  img {
    display: block;
    margin: auto;
    width: 90%;
    height: auto;
  }
</style>

@@

The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics. Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations. This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits. Outside of genome-wide studies this valuable dataset has seen little use. Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level that may provide insight into the function and evolution of genes, families and networks. Here, we present a web application and R-package through which plant molecular biologists with little to no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest. We demonstrate the utility of this tool through comprehensive analysis and identification of potential functional variation in the nuclear auxin signaling pathway.

@@

# Results  
## TIR1/AFB genes

Auxin acts by binding to receptors that in turn target co-corepressors for degradation. Auxin receptors (six in the model plant *Arabidopsis thaliana*) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009]. The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and that this rate sets the pace of development [@Guseman_Auxin-induced_2015].

Although it is unknown whether different receptors play different roles at the organismal level, a number of studies have pointed out differences in the ways the six different receptors in A. thaliana differ in their expression. The expression of AFB5 is induced in the presence of auxin, making this a circadian-regulated gene [@Covington_Circadian_2007]. Also, AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010]. Although TIR1/AFBs are expressed ubiquitously in A. thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009]. AFB1 has a mutation drastically reducing its ability to assemble into an SCF complex, contributing to its increased accumulation, as well as preventing it from degrading Aux/IAAs[@Yu_Untethering_2015]. Higher order TIR1/AFB mutants show AFB1 to have inconclusive function. However, AFB1 is shown to be able to accumulate ubiquitously in all tissues [@Parry_Complex_2009]. AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while others have a preference for naturally occurring auxin, indole-3-acetic acid [@Prigge_Arabidopsis_2016]. Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@Havens_Synthetic_2012]. TIR1 and AFB2 have slightly different speed at which they perform this function, with AFB2 being generally faster.

```{r}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "AFB_gene_ids.csv", package = "r1001genomes"))
annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

@@

Examining the natural sequence variation across the AFB family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance. *AFB3* and *AFB4* had higher nonsynonymous diversity, while their sisters genes, *AFB2* and *AFB5* were more conserved. This matches our current understanding of *AFB3* and *AFB4* as being minor players in the auxin signaling pathway, but brings in to question the inconclusive role of *AFB1* in auxin signaling.

Although most known functional regions are highly conserved in AFB1, there is a nonsynonymous polymorphism in the TIR1/AFB oligomerization domain, only found in the Can-1 accession. Mutations in this domain of TIR1 frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a]. Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.

```{r AFB1 variants for PyMOL, eval=FALSE}
coding_vcf <- ldply(.data = VCF, .fun = subset, 
             !is.na(Transcript_ID) & gt_GT != "0|0")
aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
               seq_pos = Codon_Number)
collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))

@@

```{python, eval=FALSE}
#
# -- readSelection.py
#
from pymol import cmd, stored

@@

```{pymol, eval=FALSE}
run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py

### Natural variation
```{r}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "AFB_gene_ids.csv", package = "r1001genomes"))
annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "AFBs_VCF.rds", package = "r1001genomes"))
div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
```

```{python}
#
# -- readSelection.py
#
from pymol import cmd, stored

```{python eval=FALSE}
#from within pymol
cmd.load("Dropbox/180705_NIH_renewal/Aim1-figures/2P1Q.pdb")
hide everything

@@

```{r AFB1 diversity plot}
coding_variants <- getCodingDiv(VCF[[gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]]])
p <- plotCodingDiv(uniqueCodingVars = coding_variants)
annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)
annotation
# translate positions from TIR1 to AFB1

AFB1_positions <- annotation$positions
AFB1_positions %<>% left_join(y = select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
AFB1_positions$position <- AFB1_positions$AT4G03190.1

@@

The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains. This extension had very high nonsynonymous diversity(Fig. 4), suggesting that this extension does not play an important functional role in these proteins.

Figure 4: Alignment of A. thaliana TIR1/AFB F-box domains showing variants. Cyan marks the F-box binding domain. Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively. Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
```{r TIR1/AFB Alignment}
aln_df <- makeAlnDF(alignment[[2]])
aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
aln_df$tair_symbol <- fct_rev(fct_relevel(aln_df$tair_symbol, "TIR1"))
aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
VCF <- getCodingDiv(VCF)
aln_df <- addSNPsToAlnDF(aln_df, VCF)
F_box <- subset(aln_df, aln_pos <= 96)
F_box <- chunkAlnDF(aln_df = F_box, chunk_width = 80)
p <-ggplot(F_box, aes(x = aln_pos, y = tair_symbol,
                             group = seq_pos, text = variants))
p <- p +
  geom_tile(data = na.omit(F_box), mapping = aes(fill = effects),
            width = 1, height = 1, alpha = 0.8) +
  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
  scale_x_continuous(breaks=seq(1,max(F_box$aln_pos), by = 10)) +
  scale_y_discrete() +
  # expand increases distance from axis
  labs(x = "", y = "", fill = "variant") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.box = "vertical")
p
```

## Aux/IAA genes

The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes. Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as a co-receptor to bind and perceive intracellular auxin. Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron motif [@tan_mechanism_2007]. Formation of this auxin co-receptor complex triggers ubiquitination and subsequent degradation of IAAs. This degradation in turn relieves their repression upon ARF transcription factors.Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004]. The *Arabidopsis thaliana* Aux/IAA gene family has 29 members, most of which possess a canonical degron. Notable exceptions are the auxin-insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012]. Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@moss_rate_2015; @dreher_arabidopsis_2006]. IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via hetero-oligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].

```{r, fig.height=11}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "IAA_gene_ids.csv", package = "r1001genomes"))
#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "IAAs_VCF.rds", package = "r1001genomes"))
div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
                                       .fun = function(x) 
                                         {as.integer(str_remove(x, 
                                              pattern = "IAA"))})
formatRound(datatable(div_stats[order(div_stats$tair_symbol),],
                  colnames = c("transcript",
                               "symbol",
                               "&pi;<sub>N</sub>",
                               "&pi;<sub>S</sub>",
                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
                               "&pi; coding",
                               "&pi; transcript"),
                  rownames = FALSE, escape = FALSE,
                  options = list(paging=FALSE, searching=FALSE)),
                columns = 2:7, digits = 6)
```

Within the available natural sequence variation, we found that for the canonical Aux/IAAs, no missense mutations were present in the KR motif or residues responsible for PB1-PB1 domain interactions (Supp. Fig. Alignment of canonical IAAs). The majority of nonsynonymous mutations observed were in the variable regions between the conserved domains. However, several missense polymorphisms of interest were identified in EAR domains and degrons of several canonical IAAs.

@@

Previous research has ~~pointed to~~ **found evidence of** both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005]. For example, the iaa8-1, iaa9-1 double mutant and the iaa5-1, iaa6-1, iaa19-1 triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the IAA6/19 sister pair has significant differences in expression patterns, protein abundances and functions, suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].

In our analysis of the full IAA gene family, one member of most IAA sister pairs had high nonsynonymous diversity while the other sister was more conserved (Figure IAA phylogeny heat map). In an interspecific comparison between *A. thaliana* and *A. lyrata*, IAA19 is more conserved, while IAA6 has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017]. Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for IAA19 (${\pi_N/\pi_S}$ = 0.55) compared to IAA6 (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of IAA6 as seen in Winkler et al. (Supp Fig IAA6 diversity plot). 
```{r eval=FALSE}
write.nexus.data(as.matrix(alignment[[2]]), file = "IAAs.nex", format = "protein")
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
```
```{bash eval=FALSE}
mb

execute manuscript/IAAs.nex
# set IAA33 as the outgroup
outgroup AT5G57420.1

@@

```{r}
tree <- treeio::read.beast(file = "IAAs.nex.con.tre")

@@

IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
tree@phylo$tip.label <- IAA_data$name
ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
  theme(legend.position = c(.9,.7)) + 
  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3)

@@

```{r IAA tree, fig.height=5, fig.width=5}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(hjust = -0.25) + 
  geom_tippoint(mapping = aes(size = PiN_PiS, color = PiN_PiS), show.legend = TRUE) + 
  scale_color_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
                                        max(IAA_data$PiN_PiS)), 
                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
  theme(legend.position = c(.9,.7)) + 
  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) 
p
```

>>>>

```{r setup}
knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')

#devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
library(ape)
library(r1001genomes)
library(DT)
library(tidyverse)
library(magrittr)
library(RColorBrewer)
library(ggpmisc)
library(ggthemes)
library(ggseqlogo)
library(ggtree)
library(reticulate)
library(scales)
library(viridis)
library(ips)
library(knitr)
#use_python(python = "/usr/local/bin/python")
```
<!--This is a CSS (Cascading Style Sheet).
 It controls how panels and images are formatted.></!-->
<style>
  img {
    display: block;
    margin: auto;
    width: 90%;
    height: auto;
  }
</style>

@@

The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics. Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations. This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has and is influencing the evolution of plant genomes. Outside of genome-wide studies however, this valuable dataset has seen little use. Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level that may provide insight into the function and evolution of genes, families and networks. Here, we present a web application and R-package through which plant molecular biologists with little-to-no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest. We demonstrate the utility of this tool through comprehensive analysis and identification of potential functional variation in the nuclear auxin signaling pathway.

@@

# Results  
## *TIR1/AFB* genes

Auxin acts by binding to receptors that in turn target co-repressors for degradation. Auxin receptors (six in the model plant *Arabidopsis thaliana*) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009]. The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and this rate sets the pace of lateral root development [@Guseman_Auxin-induced_2015].

Although it is unclear what unique roles each receptor plays in growth and development, a number of studies have pointed out differences in the ways the six different receptors in A. thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016]. Although TIR1/AFBs are expressed ubiquitously in A. thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009]. Additionally, the expression of AFB5 is stronly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist. 
  Functionally, all members of this family have been shown to bind auxin and Aux/IAA proteins. However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015]. This lack of complexation may allow observed high ubiquitous AFB1 accumulation [@Parry_Complex_2009]. Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling. Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@Prigge_Arabidopsis_2016]. Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @Havens_Synthetic_2012]. AFB2 induced the degradation of certain Aux/IAA proteins at a faster rate than TIR1, suggesting some functional specificity has arisen since the initial duplication between the *TIR1/AFB1* and *AFB2/AFB3* clades.

```{r TIR1_AFB_div_stats}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "AFB_gene_ids.csv", package = "r1001genomes"))
annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

@@

Examining the natural sequence variation across the *AFB* family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling. *AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved. This matches our current understanding of *AFB3* as playing a minor role in the auxin signaling pathway and *AFB4* perhaps undergoing pseudogenization.

Although most known functional regions are highly conserved in *AFB1*, there is a nonsynonymous polymorphism  in the oligomerization domain, only found in the Can-0 accession. Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a]. Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.

@@

```{python readSelection, eval=FALSE}
#
# -- readSelection.py
#
from pymol import cmd, stored

@@

```{PyMol, AFB_structure, eval=FALSE}
#from within pymol
cmd.load("Dropbox/180705_NIH_renewal/Aim1-figures/2P1Q.pdb")
hide everything

@@

```{r AFB1 diversity plot}
coding_variants <- getCodingDiv(VCF[[gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]]])
p <- plotCodingDiv(uniqueCodingVars = coding_variants)
annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)

# translate positions from TIR1 to AFB1
AFB1_positions <- annotation$positions
AFB1_positions %<>% left_join(y = select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
AFB1_positions$position <- AFB1_positions$AT4G03190.1

@@

The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains. This extension had very high nonsynonymous diversity(Fig. 4), suggesting that this extension does not play an important functional role in these proteins. Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.

Figure 4: Alignment of A. thaliana TIR1/AFB F-box domains showing variants. Cyan marks the F-box binding domain. Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively. Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
```{r TIR1/AFB F-box Alignment}
aln_df <- makeAlnDF(alignment[[2]])
aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
aln_df$tair_symbol <- fct_rev(fct_relevel(aln_df$tair_symbol, "TIR1"))
aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
VCF <- getCodingDiv(VCF)
aln_df <- addSNPsToAlnDF(aln_df, VCF)
F_box <- subset(aln_df, aln_pos <= 96)
F_box <- chunkAlnDF(aln_df = F_box, chunk_width = 80)
p <-ggplot(F_box, aes(x = aln_pos, y = tair_symbol,
                             group = seq_pos, text = variants))
p <- p +
  geom_tile(data = na.omit(F_box), mapping = aes(fill = effects),
            width = 1, height = 1, alpha = 0.8) +
  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
  scale_x_continuous(breaks=seq(1,max(F_box$aln_pos), by = 10)) +
  scale_y_discrete() +
  # expand increases distance from axis
  labs(x = "", y = "", fill = "variant") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.box = "vertical")
p
```

## *Aux/IAA* genes

The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes. Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as co-receptors to bind and perceive nuclear auxin. Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron motif [@tan_mechanism_2007]. Formation of this auxin co-receptor complex triggers ubiquitination and subsequent degradation of IAAs. This degradation in turn relieves their repression upon ARF transcription factors. Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004]. The *Arabidopsis thaliana* Aux/IAA family has 29 members, most of which possess a canonical auxin degron. Notable exceptions are the auxin-insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012]. Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@dreher_arabidopsis_2006; @moss_rate_2015]. IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via hetero-oligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].

```{r IAA diversity stats, fig.height=11}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "IAA_gene_ids.csv", package = "r1001genomes"))
#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "IAAs_VCF.rds", package = "r1001genomes"))
div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
                                       .fun = function(x) 
                                         {as.integer(str_remove(x, 
                                              pattern = "IAA"))})
div_stats <- formatRound(datatable(div_stats[order(div_stats$tair_symbol),],
                  colnames = c("transcript",
                               "symbol",
                               "&pi;<sub>N</sub>",
                               "&pi;<sub>S</sub>",
                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
                               "&pi; coding",
                               "&pi; transcript"),
                  rownames = FALSE, escape = FALSE,
                  options = list(paging=FALSE, searching=FALSE)),
                columns = 2:7, digits = 6)

kable(div_stats$x$data)
#saveWidget(div_stats, file = "IAA_div_stats.html", )

Within the available natural sequence variation, we found several missense polymorphisms affecting the EAR domain or degron of several canonical IAAs. However, no missense polymorphisms were present in the KR motif or residues critical for PB1-PB1 domain interactions (Supp. Fig. Alignment of canonical IAAs). The majority of missense mutations observed were in the variable regions between the conserved domains.

@@

Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005]. For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].

In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved (Figure IAA phylogeny heat map). In an interspecific comparison between *A. thaliana* and *A. lyrata*, *IAA19* is more conserved, while *IAA6* has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017]. Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of *IAA6* as seen in Winkler et al. (Supp Fig IAA6 diversity plot). 
```{r IAA alignment}
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

# make a collapsed alignment for translating positions
collapsed_aln_df <- 
  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
  tidyr::spread(key = seq_name, value = seq_pos)
```

```{r Mask_IAA_alignment, eval=FALSE}
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
```
```{bash IAA_phylogeny_construction, eval=FALSE}
~/Desktop/natural-variation-webtool/manuscript

/Applications/MrBayes/mb

execute IAAs.nex
# set IAA33 as the outgroup
outgroup AT5G57420.1

@@

```{r Load IAA phylogeny and data}
tree <- treeio::read.beast(file = "IAAs.nex.con.tre")

@@

IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
tree@phylo$tip.label <- IAA_data$name
```
*I think IAA20, 30, and 15 should also be in class C. So there are really two large clades A and B and then there are truncations/loss of function/dominant variants scattered throughout, Class C.*

```{r IAA_tree, fig.height=5, fig.width=5, fig.cap="(ref:IAA tree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
  geom_tippoint(mapping = aes(size = 4*PiN_PiS, fill = PiN_PiS), shape = 21, color = "white", show.legend = TRUE) + 
  scale_fill_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
                                        max(IAA_data$PiN_PiS)), 
                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
  theme(legend.position = c(.9,.7)) + 
  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
  geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
  scale_color_viridis(discrete = TRUE)
p
```

(ref:IAA tree) IAA protein sequence tree mapped with ${\pi_N/\pi_S}$

Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(IAA tree) [@remington_contrasting_2004]. Interestingly, among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity. The *Arabidopsis thaliana* EFP browser shows that **IAA10** and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**. In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of *IAA10* and *IAA11* playing redundant roles. Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as *IAA10* and *IAA11*.

### New Loss-of-Function/Gain-of-Function alleles
Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs. For example, three accessions have a missense variant affecting the EAR domain of IAA10 (HR-5, UKSE06-118 and UKSW06-240, all from southern UK). This variant (L49S) is in the second leucine of the LxLxL motif and likely reduces the ability of IAA10 to recruit the TOPLESS co-repressor [@lee_biological_2016]. The nonsense variant (IAA10-Y219\*) resulting in truncated PB1 domain may be more indicative of *IAA10* null phenotype, as it likely prevents IAA10 from interacting with an ARF. This variant is only found in one accession from northeastern Spain (IP-Svi-0). These accessions and potential loss-of-function alleles may facilitate exploration of the function of *IAA10*.

Several examples of possible gain-of-function missense variants were found in the degron of *IAA2*, *IAA3*, *IAA6*, *IAA7*, *IAA8*, *IAA13*, *IAA15*, *IAA18*, *IAA19*, *IAA27*, *IAA28* and *IAA29*. Most degron variants were very minor alleles in the population, only represented in one or two accessions. Interestingly, degron missense variants occured more frequently in *IAA* genes that have a closely related sister, which may indicate redundant function within the sister-pairs. *I think this is an interesting finding, but I'm not sure about this hypothesis of redundant function. If pairs are acting redundantly and we assume the degron variants are dominant, we would expect the auxin signal they transmit (their function) to be lower, in sum, as a pair. So I would actually hypothesize that there has been some divergence in function as a pair. * 
```{r Correlation between Distance to MRCA and prevalence of degron variants}

@@

```{r Correlation between Pi_N/Pi_S and k_5}

*IAA6* has two potentially dominant missense mutations found within the degron (Supp Fig. alignment), one previously believed to lead to the *shy1-1* phenotype [@kim_two_1996]. Interestingly, this variant is widely represented in 31 accessions across Europe (Supplemental Map Fig). This finding supports the proposition that iaa6-C78R may not be the mutation causing the *shy1-1D* phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)). These accessions could be used to confirm whether *iaa6-C78R* is responsible for the *shy1-1D* phenotype. Additionally, an accession was identified with the IAA6-V77M variant, also in the degron, that may be useful to further explore *iaa6* gain-of-function phenotypes.
